\chapter{Advanced substitution}

This chapter includes some cases of substitution that require little bit more thinking than normal substitute commands.

\section{Rearrange CSV fields using submatches}

You can use submatches in search to rearrange one or more words.\\
\verb|:%s/{search pattern with submatches}/\{submatch 1-9}|
See pg 228 in Practical Vim for a detailed example.

\section{Perform arithmetic on the replacement}

You can perform arithmetics on the matches and replace the old matches with the result.
For example, \verb|:%s//\=submatch(0)-1/g| will minus 1 to whatever matches in the search.
\verb|submatch({number})| returns the \verb|{number}|'th submatch of the matched text.
When 0 is used, the whole matches text is returned.
\verb|submatch()| can only be used in the \verb|:s| command or the \verb|substitute()| function.
Another thing to note is that you have to use this function to access the submatches in vim script.

\section{Swap two or more words}

You can use a dictionary data type for this \verb|:let x={"dog":"man", "man":"dog"}|\\
You can access it by \verb|x["key"]| (e.g. \verb|:echo x["dog"]| will give man).
You also need to be able to match the words that you want to swap.\\
\verb-/\v(<"word1">|<"word2">...)- will capture the match into the submatches.
To swap  the words, type:\\
\verb|:%s//\={dictionary of words to be swapped}[submatch(1)]/g|\\
This uses the submatch to look up the matching keyword in the dictionary you have defined, and use the other key value as the replacement string.
As a result, you would have swapped the words.
This approach will allow you to swap 2 or more words in a document with a single hit.
(checkout the plugin discussed on pg 232 in Practical Vim).

\section{Find and replace across multiple files}

You need to first set up the substitute command, then use the argument list and \verb|:argdo| command to apply it to all the files.
You can use the \verb|e| flag to suppress the error messages when it can't find any matches in a file.
This method will attempt to substitute in all  the files, whether or not it has the match (see pg 235-236 for an alternative solution).
